Information Closure Theory of Consciousness

### 2020-05-15 ICT_Frontiers_rev202001.tex

	\begin{abstract}
		Information processing in neural systems can be described and analysed at multiple spatiotemporal scales. Generally, information at lower levels is more fine-grained and can be coarse-grained in higher levels. However, information processed only at specific scales of coarse-graining seems to be available for conscious awareness. We do not have direct experience of information available at the scale of individual neurons, which is noisy and highly stochastic. Neither do we have experience of more macro-scale interactions such as interpersonal communications. Neurophysiological evidence suggests that conscious experiences co-vary with information encoded in coarse-grained neural states such as the firing pattern of a population of neurons. In this article, we introduce a new informational theory of consciousness: \acf{OurTheory}. We hypothesise that conscious processes are processes which form \ac{ntic} with respect to the environment at certain coarse-grained scales. This hypothesis implies that conscious experience is confined due to informational closure from conscious processing to other coarse-grained scales. \ac{OurTheory} proposes new quantitative definitions of both conscious content and conscious level. With the parsimonious definitions and a hypothesise, \ac{OurTheory} provides explanations and predictions of various phenomena associated with consciousness. The implications of \ac{OurTheory} naturally reconciles issues in many existing theories of consciousness and provides explanations for many of our intuitions about consciousness. Most importantly, \ac{OurTheory} demonstrates that information can be the common language between consciousness and physical reality.

@@

		%% Story about a cell
		Imagine you are a neuron in Alice's brain. Your daily work is to collect neurotransmitters through dendrites from other neurons, accumulate membrane potential, and finally send signals to other neurons through action potentials along axons. However, you have no idea that you are one of the neurons in Alice's supplementary motor area and involved in many motor control processes for Alice's actions, for example, grabbing a cup. You are ignorant of intentions, goals, and motor plans that Alice has at every moment even though you are part of the physiological substrate responsible for all those actions.
		A similar story also happens to Alice's conscious mind. To grab a cup, for example, Alice is conscious of her intention and visuosensory experience of this action. However, her conscious experience does not reflect the dynamic of your membrane potential or the action potentials you send to other neurons every second. That is, not all information you have is available to Alice's conscious mind.

@@

		It seems to be true that we are not consciously accessing information processed at every scale in the neural system. There are both more microscopic and more macroscopic scales than the scale corresponding to the conscious contents. On the one hand, dynamics of individual neurons are stochastic \citep{Goldwyn2011, White2000}. However, what we are aware of in our conscious mind shows astonishing stability and robustness against the ubiquitous noise in the neural system \citep{mathis1995computational}. In addition, some parts of the neural system contribute very little to conscious experience (the cerebellum for example \citep{lemon2010life}), also suggesting that conscious contents do not have one-to-one mapping to the entire state of the neural system. On the other hand, human conscious experience is more detailed than just a simple (e.g. binary) process can represent, suggesting that the state space of conscious experience is much larger than what a single overly coarse-grained binary variable can represent. These facts suggest that conscious processes occur at a particular scale. \linelabel{line:lack-theory}We currently have only few theories (e.g., Integrated Information Theory \citep{hoel2016can} and Geometric Theory of Consciousness \citep{fekete2011towards,fekete2012lack}) to identify the scale which conscious processes correspond to ( also see discussion in \cite{fekete2016system}). We refer to this notion as \textbf{the scale problem of consciousness} (Fig.~\ref{fig:scaleproblem}).

		%% our argument
		In this article, we propose a new information-based theory of consciousness, called \acf{OurTheory}. We argue that every process with a positive non-trivial information closure (NTIC) has consciousness. This means that the state of such a process corresponds one-to-one to conscious content.\footnote{In the following IC stands for "informational closure" or "informationally closed" and NTIC stands for "non-trivial informational closure" or "non-trivially informationally closed".}. We further postulate that the \textit{level} of consciousness corresponds to the degree of NTIC. (for a discussion of the distinction between level versus content of consciousness see \cite{laureys2005neural, overgaard2010neural}).

		In the following, we first introduce non-trivial informational closure and argue for its importance to information processing for human scale agents (Sec.~\ref{sec:Non-trivial informational closure}). We next argue that through coarse-graining the neural system can form informational closure and a high degree of NTIC at a specific scale of coarse-graining  (Sec.~\ref{sec:Neural coarse-graining}). In the Sec.~\ref{sec:OurTheory}, we propose a new theory of consciousness (\ac{OurTheory}). We also illustrate how \ac{OurTheory} can parsimoniously explain empirical findings from previous consciousness studies (Sec.~\ref{sec:Conscious versus Unconscious Processing}) and reconcile several current major theories of consciousness (Sec.\ref{sec:Comparison with other theories}). Finally, we discuss the current theoretical and empirical limitations of \ac{OurTheory} and propose the implications from \ac{OurTheory} to the current consciousness science (Sec.\ref{sec:Limitation and Future work}).

@@

% ============================================================================ %
%                       Non-trivial informational closure                      %
% ============================================================================ %
	\section{Non-trivial Informational Closure} \label{sec:Non-trivial informational closure}
		The notion of non-trivial informational closure (NTIC) is introduced by \cite{BERTSCHINGER.2006}. The concept of closure is closely related to system identification in systems theory. One can distinguish a system from its environment by computing the closedness of the system \citep{maturana1991autopoiesis, rosen1991life, pattee2012evolving, luhmann1995probleme}. The closedness can be further quantified by information theory.

		% Definition of informational closure
			\linelabel{line:r1q1a}Consider two processes, the environment process $(E_t)_{t \in \mathbb{N}}$ and the system's process $(Y_t)_{t \in \mathbb{N}}$ and let their interaction be described by the Bayesian network with the sensor channel $\hat{e}_{t}$ and the action $\hat{y}_{t}$ channel in Fig.~\ref{fig:SystemAndEnv}. Then, information flow $J_{t}$ from the environment $E$ to a system $S$ at time $t$ can be defined as the conditional mutual information $I$ between the current environment state $E_{t}$  and the future system state $Y_{t+1}$ given the current system state $Y_{t}$

@@

				\begin{figure}
					\includegraphics[width=\textwidth]{Submission/Frontiers/SystemAndEnv_3.pdf}
					\caption{The dependencies between a system  $Y$ and its environment $E$ through the channels $\hat{y_t}$ and $\hat{e_t}$.} % Figure adapted from \cite{BERTSCHINGER.2006}
					\label{fig:SystemAndEnv}
				\end{figure}

			\noindent
			\cite{BERTSCHINGER.2006} defines that a system is informationally closed when information flow from the environment to the system is zero.

@@

			\noindent
			This implies the system contains in itself all the information about its own future and the self-predictive information contains the information about the environment.

			\subsection{Informational Closure Does not Imply Causality}\label{sec:causality}
                A surprising result from the definition of information flow $J_t(E\rightarrow Y)$ (Eq.~\ref{eq:InformationFlow}) is that information flow does not indicate causal dependency from $E_t$ to $Y_{t+1}$ and from $Y_t$ to $Y_{t+1}$. Here we consider two scenarios, called \textit{modelling} and \textit{passive adaptation}, which were already noted by \citet{BERTSCHINGER.2006}. In both scenarios, a process can form positive NTIC ($NTIC(E\rightarrow Y)> 0$) and informational closure ($J(E\rightarrow Y)=0$), however, via different causal dependencies.

@@

                In the scenario of \textit{passive adaptation}, counterintuitively, the future system states ($Y_{t+1}$) are  entirely driven by the current environment states ($E_t$) but can also have positive NTIC and informational closure. This happens under the condition that the sensory process $\hat{e}_t$ is deterministic and the system simply copies the sensory values. The system is then a copy of another informationally closed process ($\hat{e}_t$) and is therefore closed itself. At the same time it has mutual information with the process that it is copying.

                However, in most of the realistic cases, the environment is partially observable to the system and, thus, the sensory process is usually not deterministic. Therefore, it is difficult for the system to be informationally closed and have higher NTIC.
                More importantly, we argue in Appendix that whenever the environment has itself more predictable dynamics than the observations, there is potential for a process to achieve higher NTIC by modelling the environment than by copying the observations.

@@

		%% stochasticity at microscopic levels
		The formation of NTIC with a highly stochastic process is challenging. NTIC requires the predictability of the system state and is therefore impeded by noise in the system. Information processing at the microscopic scale (the cellular scale) in neural systems suffers from multiple environmental noise sources such as sensor, cellular, electrical, and synaptic noises. For example, neurons exhibit large trial-to-trial variability at the cellular scale, and are subject to thermal fluctuations and other physical noises \citep{faisal2008noise}.

        %% Our argument
		However, it is possible that neural systems form NTIC at certain macroscopic scales through coarse-graining of microscopic neural states. Coarse-graining refers to many-to-one or one-to-one maps which aggregate microscopic states to a macroscopic state. In other words, a number of different micro-states correspond to the same value of the macro-variable \citep{price2007causation}. Coarse-grainings, can therefore form more stable and deterministic state transitions and more often form NTIC processes. For neural systems this means that a microscopically noisy neural system may still give rise to an NTIC process on a more macroscopic scale.

		Indeed, empirical evidence suggests that coarse-graining is a common coding strategy to establish robustness against noise at microscopic scales of the neural system. For instance, the inter-spike intervals of an individual neuron are stochastic. This implies that the state of an individual neuron does not represent stable information. However, the firing rate, i.e. the average spike counts over a given time interval, is more stable and robust against noise such as the variability in inter-spike intervals. Using this temporal coarse-graining strategy, known as rate coding \citep{adrian1926impulses, gerstner2002spiking, maass2001pulsed, panzeri2015neural, stein2005neuronal}, neurons can encode stimulus intensity by increasing or decreasing the firing rate \citep{kandel2000principles}. \citep{stein2005neuronal}. The robustness of the rate coding is a direct consequence of the many to one mapping (i.e., coarse-graining).

		% population code
		Population coding is another example of encoding information through coarse-graining in neural systems. In this coding scheme, information is encoded by activation patterns of a set of neurons (a neuron population). In the population coding scheme, many states for a neuron population map to the same state of macroscopic variables which encode particular informational contents, thereby reducing the influence of noise in individual neurons. That is, stable representations can be formed through coarse-graining the high dimensional state space of a neuron population to a lower dimensional macroscopic state space \citep{kristan1997population, pouget2000information, binder2009encyclopedia, QuianQuiroga2009}. Therefore, individual neuron states (the microscopic scale) are not informative enough about the complete encoded contents at the population scale (the macroscopic scale). Instead, coarse-grained variables are better substrates for stably encoding information and allow the neural system to ignore noisy interactions at the fine-grained scale \citep{Woodward2007-WOOCWA}.

        These two examples show that the known coding schemes can be viewed as coarse-graining, and provide stochastic neural systems with the ability to form more stable and deterministic macroscopic processes for encoding and processing information reliably. We argue that through coarse-graining the neural systems is able to form NTIC processes at macroscopic scales. Based on the merit of coarse-graining in neural systems, we propose a new theory of consciousness in the next section.

% ============================================================================ %
%               A neural coarse graining theory of consciousness               %
% ============================================================================ %
	\section{\acf{OurTheory}}\label{sec:OurTheory}

        In this section, we propose a new theoretical framework of consciousness: \acf{OurTheory}. The main hypothesis is that conscious processes are captured by what we call \textit{C-processes}. We first define C-processes, then state our hypothesis and discuss its implications.

@@

	   	% ------------------------------- definition 1 ------------------------------- %
	   	In order to define C-processes we need to define coarse-grainings first. Every coarse-graining is characterised by a function that maps the microscopic process to the coarse-grained macroscopic process. More formally:

	   	\begin{definition}
	   	\linelabel{line:r2q1-1} Given a stochastic process $X$ with state space $\mathcal{X}$, a \emph{coarse-graining of $X$} is a stochastic process $Y$ with state space $\mathcal{Y}$ such that there exists a function \footnote{Functions in the mathematical sense used here are always either one-to-one or many-to-one.} $f_Y:\mathcal{X}\rightarrow \mathcal{Y}$ with $Y_t=f_Y(X_t)$.  
	   	\end{definition}
	   	A more general definition of coarse-grainings that maps temporally extended sequences of the microscopic process to macroscopic states are possible but for this first exposure of our theory the simpler definition above is sufficient.

@@

        \noindent Note that, \linelabel{line:IC2CG}here we applied the same definitions of information flow (Eq.~\ref{eq:InformationFlow})
        \begin{align}
            J_{t}(E \rightarrow Y )=I(Y_{t+1};E_t|Y_t)
        \end{align}
            to the system-environment dependency
            and the micro-macro scale dependency
        \begin{align}
            J_{t}(X \rightarrow Y )=I(Y_{t+1};X_t|Y_t)
        \end{align}
        Even though the Bayesian graphs are different in the two scenarios. Both of these settings have already been used in the literature \citep[see][]{BERTSCHINGER.2006, PFANTE.2014}).\\

@@

            \item[Implication 3.] 
            Conscious processes are self-determining. This is a direct implication of the requirement that $Y$ is informationally closed with respect to $X$. To be informationally closed with respect to $X$, no coarse-graining knows anything about the conscious process' future that the conscious process does not know itself. This self-determining characteristics is also in line with our daily life conscious experience which often shows stability and continuity and is ignorant of the stochasticity (e.g., noise) of the cellular scales.

@@

            \item[Implication 5.]
            Conscious processes can model environmental information (by forming NTIC), however, be ignorant to part of the information of more microscopic processes (from Implication 3 and 4). This is in line with our conscious experience that information that every conscious percept provides represent rich and structured environmental states without involving all the information about microscopic activities.
        \end{description}

@@

            This implies that a highly NTIC process must have rich dynamics with self-predictability over time. Another implication is that complex systems can potentially attain higher levels of consciousness due to the larger information capacities needed to attain high mutual information. This outcome is consistent with the common intuition that conscious levels are often associated with the degree of complexity of a system.

@@

    	    %% not monotonic
    		\linelabel{line:non-monotonic}It is important to note that NTIC can be a non-monotonic function of the scale of coarse-graining. At the finest scale we consider the whole universe $X$ as the process $Y$. then, since $Y$ is a coarse graining of $S$ we have $Y=S=X$. In this case the environment $E$ corresponding to the universe seen as a system is the constant coarse-graining\footnote{Recall that, for a system with state space $\S$ the environment state space $\E$ must be such that $\X=\S \times \E$. If $\S =\X$ then we need $\E$ with $\X \times \E=\X$ such that $\E$ must be a singleton set. All coarse-grainings mapping $\X$ to a singleton set are constant over $\X$.} and therefore the mutual information $I(E_t;Y_{t+1})$ and the transfer entropy $I(Y_{t+1};E_t|Y_t)$ are zero. Then NTIC of the universe with respect to its environment is zero and $X$ can never be a C-process.

    		If we now increase the scale of $Y$, this allows $S$ to also reduce in scale and therefore $E$ can become more and more fine-grained. This means that the mutual information $I(E_t;Y_{t+1})$ between $E$ and $Y$ can at least potentially become positive. Up to the point where $E$ accounts for half of the bits of $X$ and $S$ for the other half the upper bound of the mutual information $I(E_t;Y_{t+1})$ achieved when $Y=S$ increases. Even further refining $E$ then again leads to a reduction of the upper bound of $I(E_t;Y_{t+1})$.
            % ==================================== %
            %    WHY DO WE NEED the FUCKING $Y$    %
            % ==================================== %
            % 2020/04/07 01:57
    		% Why does Definition 2 have a fucking $Y$? It seems the whole thing could be defined wihtout referring to the underlying $S$ and directly use $Y$'s own environment (together with which it corresponds to $X$). We would always have a refinement of $Y$ that could be called $S$ to gether with its environment. But this is alwyas the case so it doesn't restrict the definition of $Y$ at all. (Especially if we can just choose $S=Y$). The only thing that makes the current definition of $Y$ more general is that it may be that $Y$ is only closed with respect to a coarsening of its own environment and not to its own environment itself. This is a waekening of the restriction compared to requiring that $Y$ be NTIC with respect to its own environment and IC with respect to $X$. Actually no, since $Y$ is IC with respect $X$ it is IC with respect to everything including to any coarsening of its own environment. Since the mutual information between $Y$ and its environment is greater or equal to the mutual information between $Y$ and a coarsening of its environment $Y$ must also be NTIC with respect to its own environment. 
    		% ==================================== %

    		On the other extreme, when $E=X$ the system state space must be the singleton set and NTIC from $E$ to $Y$ must again be zero. Therefore, processes at intermediate scales of coarse-graining can form higher degrees of NTIC than at the most microscopic or macroscopic scales (Fig.~\ref{fig:LevelOfConsciousness}). \ac{OurTheory} suggests that human consciousness occurs at a scale of coarse-graining where high NTIC is formed within the neural system. \linelabel{line:non-monotonic_end}

    		\begin{figure}[H]				
        		\includegraphics[width=\textwidth]{Submission/Frontiers/nonMono.pdf}
        		\caption{A non-monotonic relationship between scale of coarse-graining and level of consciousness.}
        		\label{fig:LevelOfConsciousness}
    		\end{figure}

		\subsection{Conscious Contents Corresponding to States of a NTIC Process}\label{sec:cc}
    	    % richness 
    		\ac{OurTheory} proposes that conscious contents correspond to the states of NTIC processes (Eq.~\ref{eq:cContent}). This implies that the size of the state space of an NTIC process is associated with the richness of conscious contents that the process can potentially have. Therefore, a complex NTIC process with a high dimensional state space can have richer conscious experience than a simple NTIC process can have. This outcome is consistent with the intuition that the richness of conscious contents is associated with the complexity of a system.

    		% CC doesn't include all microscopic states
    		Informational closure can happen between scales of coarse-graining within a single system. Thus, a macroscopic NTIC process can be ignorant to its microscopic states. \ac{OurTheory} argues that human conscious contents do not reflect cellular scale activity because the conscious process which corresponds to a macroscopic NTIC process is informationally closed to the cellular scale in the human neural system. Further more, since NTIC processes are informationally closed, each of them can be considered as a reality. When the information flow from its microscopic processes and environment to the informationally closed process is completely zero (Eq.~\ref{eq:informationflow2}), the future states of the process can entirely be determined by its past states.

>>>>

	\begin{abstract}
		Information processing in neural systems can be described and analysed at multiple spatiotemporal scales. Generally, information at lower levels is more fine-grained but can be coarse-grained at higher levels. However, only information processed at specific scales of coarse-graining appears to be available for conscious awareness. We do not have direct experience of information available at the scale of individual neurons, which is noisy and highly stochastic. Neither do we have experience of more macro-scale interactions, such as interpersonal communications. Neurophysiological evidence suggests that conscious experiences co-vary with information encoded in coarse-grained neural states such as the firing pattern of a population of neurons. In this article, we introduce a new informational theory of consciousness: \acf{OurTheory}. We hypothesise that conscious processes are processes which form \ac{ntic} with respect to the environment at certain coarse-grained scales. This hypothesis implies that conscious experience is confined due to informational closure from conscious processing to other coarse-grained scales. \ac{OurTheory} proposes new quantitative definitions of both conscious content and conscious level. With the parsimonious definitions and a hypothesise, \ac{OurTheory} provides explanations and predictions of various phenomena associated with consciousness. The implications of \ac{OurTheory} naturally reconcile issues in many existing theories of consciousness and provides explanations for many of our intuitions about consciousness. Most importantly, \ac{OurTheory} demonstrates that information can be the common language between consciousness and physical reality.

@@

		%% Story about a cell
		Imagine you are a neuron in Alice’s brain. Your daily work is to collect neurotransmitters through dendrites from other neurons, accumulate membrane potential, and finally send signals to other neurons through action potentials along axons. However, you have no idea that you are one of the neurons in Alice’s supplementary motor area and are involved in many motor control processes for Alice’s actions, such as grabbing a cup. You are ignorant of intentions, goals, and motor plans that Alice has at any moment, even though you are part of the physiological substrate responsible for all these actions. A similar story also happens in Alice’s conscious mind. To grab a cup, for example, Alice is conscious of her intention and visuosensory experience of this action. However, her conscious experience does not reflect the dynamic of your membrane potential or the action potentials you send to other neurons every second. That is, not all the information you have is available to Alice’s conscious mind.

@@

		It appears to be true that we do not consciously access information processed at every scale in the neural system. There are both more microscopic and more macroscopic scales than the scale corresponding to the conscious contents. On the one hand, the dynamics of individual neurons are stochastic \citep{Goldwyn2011, White2000}. However, what we are aware of in our conscious mind shows astonishing stability and robustness against the ubiquitous noise in the neural system \citep{mathis1995computational}. In addition, some parts of the neural system contribute very little to conscious experience (the cerebellum for example \citep{lemon2010life}), also suggesting that conscious contents do not have one-to-one mapping to the entire state of the neural system. On the other hand, human conscious experience is more detailed than just a simple (e.g. binary) process can represent, suggesting that the state space of conscious experience is much larger than what a single overly coarse-grained binary variable can represent. These facts suggest that conscious processes occur at a particular scale. \linelabel{line:lack-theory}We currently have possess only a few theories (e.g., Integrated Information Theory \citep{hoel2016can} and Geometric Theory of Consciousness \citep{fekete2011towards,fekete2012lack}) to identify the scale to which conscious processes correspond (also see discussion in \cite{fekete2016system}). We refer to this notion as \textbf{the scale problem of consciousness} (Fig.~\ref{fig:scaleproblem}).

		%% our argument
		In this article, we propose a new information-based theory of consciousness, called the  \acf{OurTheory}. We argue that every process with a positive non-trivial information closure (NTIC) has consciousness. This means that the state of such a process corresponds one-to-one to conscious content.\footnote{In the following IC stands for "informational closure" or "informationally closed" and NTIC stands for "non-trivial informational closure" or "non-trivially informationally closed".}. We further postulate that the \textit{level} of consciousness corresponds to the degree of NTIC. (For a discussion of the distinction between level versus content of consciousness see \cite{laureys2005neural, overgaard2010neural}).

		In the following, we first introduce non-trivial informational closure and argue for its importance to information processing for human scale agents (Sec.~\ref{sec:Non-trivial informational closure}). We next argue that through coarse-graining the neural system can form informational closure and a high degree of NTIC at a specific scale of coarse-graining  (Sec.~\ref{sec:Neural coarse-graining}). In Sec.~\ref{sec:OurTheory}, we propose a new theory of consciousness (\ac{OurTheory}). We also illustrate how \ac{OurTheory} can parsimoniously explain empirical findings from previous consciousness studies (Sec.~\ref{sec:Conscious versus Unconscious Processing}) and reconcile several current major theories of consciousness (Sec.\ref{sec:Comparison with other theories}). Finally, we discuss the current theoretical and empirical limitations of \ac{OurTheory} and propose the implications of \ac{OurTheory} on the current consciousness science (Sec.\ref{sec:Limitation and Future work}).

@@

% ============================================================================ %
%                       Non-trivial informational closure                      %
% ============================================================================ %
	\section{Non-trivial Informational Closure} \label{sec:Non-trivial informational closure}
		The notion of non-trivial informational closure (NTIC) was introduced by \cite{BERTSCHINGER.2006}. The concept of closure is closely related to system identification in systems theory. One can distinguish a system from its environment by computing the closedness of the system \citep{maturana1991autopoiesis, rosen1991life, pattee2012evolving, luhmann1995probleme}. Closedness itself can be further quantified by information theory.

		% Definition of informational closure
			\linelabel{line:r1q1a}Consider two processes, the environment process $(E_t)_{t \in \mathbb{N}}$ and the system process $(Y_t)_{t \in \mathbb{N}}$ and let their interaction be described by the Bayesian network with the sensor channel $\hat{e}_{t}$ and the action $\hat{y}_{t}$ channel in Fig.~\ref{fig:SystemAndEnv}. Information flow $J_{t}$ from the environment $E$ to a system $S$ at time $t$ can then be defined as the conditional mutual information $I$ between the current environment state $E_{t}$  and the future system state $Y_{t+1}$ given the current system state $Y_{t}$

@@

				\begin{figure}
					\includegraphics[width=\textwidth]{Submission/Frontiers/SystemAndEnv_3.pdf}
					\caption{Dependencies between a system  $Y$ and its environment $E$ through the channels $\hat{y_t}$ and $\hat{e_t}$.} % Figure adapted from \cite{BERTSCHINGER.2006}
					\label{fig:SystemAndEnv}
				\end{figure}

			\noindent
			\cite{BERTSCHINGER.2006} defines a system as informationally closed when information flow from the environment to the system is zero.

@@

			\noindent
			This implies that the system contains within itself all the information about its own future and the self-predictive information contains the information about the environment.

			\subsection{Informational Closure Does not Imply Causality}\label{sec:causality}
                A surprising result from the definition of information flow $J_t(E\rightarrow Y)$ (Eq.~\ref{eq:InformationFlow}) is that information flow does not indicate causal dependency from $E_t$ to $Y_{t+1}$ or from $Y_t$ to $Y_{t+1}$. Here we consider two scenarios, called \textit{modelling} and \textit{passive adaptation}, which previously noted by \citet{BERTSCHINGER.2006}. In both scenarios, a process can form positive NTIC ($NTIC(E\rightarrow Y)> 0$) and informational closure ($J(E\rightarrow Y)=0$), albeit via different causal dependencies.

@@

                In the \textit{passive adaptation} scenario, counterintuitively, the future system states ($Y_{t+1}$) are  entirely driven by the current environment states ($E_t$) but can also have positive NTIC and informational closure. This happens under the condition that the sensory process $\hat{e}_t$ is deterministic and the system simply copies the sensory values. The system is then a copy of another informationally closed process ($\hat{e}_t$) and is therefore closed itself. At the same time it has mutual information with the process that it is copying.

                In most of the realistic cases, however, the environment is partially observable to the system, and thus the sensory process is usually not deterministic. Accordingly, it is difficult for the system to be informationally closed and have higher NTIC.
                More importantly, we argue in the Appendix that whenever the environment has itself more predictable dynamics than the observations, there is the potential exists for a process to achieve higher NTIC by modelling the environment rather than by copying the observations.

@@

		%% stochasticity at microscopic levels
		The formation of NTIC with a highly stochastic process is challenging. NTIC requires the predictability of the system state and is therefore impeded by noise in the system. Information processing at the microscopic scale (cellular scale) in neural systems suffers from multiple environmental noise sources such as sensor, cellular, electrical, and synaptic noises. For example, neurons exhibit large trial-to-trial variability at the cellular scale, and are subject to thermal fluctuations and other physical noises \citep{faisal2008noise}.

        %% Our argument
		Nevertheless, it is possible that neural systems form NTIC at certain macroscopic scales through coarse-graining of microscopic neural states. Coarse-graining refers to many-to-one or one-to-one maps which aggregate microscopic states to a macroscopic state. In other words, a number of different micro-states correspond to the same value of the macro-variable \citep{price2007causation}. Coarse-grainings can therefore form more stable and deterministic state transitions and more often form NTIC processes. For neural systems this means that a microscopically noisy neural system may still give rise to an NTIC process on a more macroscopic scale.

		Indeed, empirical evidence suggests that coarse-graining is a common coding strategy of the neural system by which it establishes robustness against noise at microscopic scales. For instance, the inter-spike intervals of an individual neuron are stochastic. This implies that the state of an individual neuron does not represent stable information. However, the firing rate, i.e. the average spike counts over a given time interval, is more stable and robust against noise such as the variability in inter-spike intervals. Using this temporal coarse-graining strategy, known as rate coding \citep{adrian1926impulses, gerstner2002spiking, maass2001pulsed, panzeri2015neural, stein2005neuronal}, neurons can encode stimulus intensity by increasing or decreasing their firing rate \citep{kandel2000principles}. \citep{stein2005neuronal}. The robustness of rate coding is a direct consequence of the many-to-one mapping (i.e., coarse-graining).

		% population code
		Population coding is another example of encoding information through coarse-graining in neural systems. In this coding scheme, information is encoded by the activation patterns of a set of neurons (a neuron population). In the population coding scheme, many states of a neuron population map to the same state of macroscopic variables which encode particular informational contents, thereby reducing the influence of noise in individual neurons. That is, stable representations can be formed through coarse-graining the high dimensional state space of a neuron population to a lower dimensional macroscopic state space \citep{kristan1997population, pouget2000information, binder2009encyclopedia, QuianQuiroga2009}. Therefore, individual neuron states (microscopic scale) are not sufficiently informative about the complete encoded contents at the population scale (macroscopic scale). Instead, coarse-grained variables are better substrates for stably encoding information and allow the neural system to ignore noisy interactions at the fine-grained scale \citep{Woodward2007-WOOCWA}.

        These two examples show that the known coding schemes can be viewed as coarse-graining, and provide stochastic neural systems with the ability to form more stable and deterministic macroscopic processes for encoding and processing information reliably. We argue that coarse-graining allows neural systems to form NTIC processes at macroscopic scales. Based on the merit of coarse-graining in neural systems, we propose a new theory of consciousness in the next section.

% ============================================================================ %
%               A neural coarse graining theory of consciousness               %
% ============================================================================ %
	\section{\acl{OurTheory}}\label{sec:OurTheory}

        In this section, we propose a new theoretical framework of consciousness: the \acf{OurTheory}. The main hypothesis is that conscious processes are captured by what we call \textit{C-processes}. We first define C-processes, then state our hypothesis and discuss its implications.

@@

	   	% ------------------------------- definition 1 ------------------------------- %
	   	To define C-processes we first need to define coarse-grainings. Every coarse-graining is characterised by a function that maps the microscopic process to the coarse-grained macroscopic process. More formally:

	   	\begin{definition}
	   	\linelabel{line:r2q1-1} Given a stochastic process $X$ with state space $\mathcal{X}$, a \emph{coarse-graining of $X$} is a stochastic process $Y$ with state space $\mathcal{Y}$ such that there exists a function \footnote{Functions in the mathematical sense used here are always either one-to-one or many-to-one.} $f_Y:\mathcal{X}\rightarrow \mathcal{Y}$ with $Y_t=f_Y(X_t)$.  
	   	\end{definition}
	   	A more general definition of coarse-grainings that maps temporally extended sequences of the microscopic process to macroscopic states are possible, but for this first exposure of our theory the simpler definition above is sufficient.

@@

        \noindent Note that, \linelabel{line:IC2CG}here we applied the same definitions of information flow (Eq.~\ref{eq:InformationFlow})
        \begin{align}
            J_{t}(E \rightarrow Y )=I(Y_{t+1};E_t|Y_t)
        \end{align}
            to the system-environment dependency
            and the micro-macro scale dependency
        \begin{align}
            J_{t}(X \rightarrow Y )=I(Y_{t+1};X_t|Y_t)
        \end{align}
        even though the Bayesian graphs differ in the two scenarios. Both these settings have been previously used in the literature \citep[see][]{BERTSCHINGER.2006, PFANTE.2014}.\\

@@

            \item[Implication 3.] 
            Conscious processes are self-determining. This is a direct implication of the requirement that $Y$ is informationally closed with respect to $X$. To be informationally closed with respect to $X$, no coarse-graining knows anything about the conscious process' future that the conscious process does not know itself. This self-determining characteristics is also consistent with our daily life conscious experience which often shows stability and continuity and is ignorant of the stochasticity (e.g., noise) of the cellular scales.

@@

            \item[Implication 5.]
            Conscious processes can model environmental information (by forming NTIC) but be ignorant to part of the information of more microscopic processes (from Implication 3 and 4). This is consistent with our conscious experience, namely that the information that every conscious percept provides represents rich and structured environmental states without involving all the information about microscopic activities.
        \end{description}

@@

            This implies that a highly NTIC process must have rich dynamics with self-predictability over time. Another implication is that complex systems can potentially attain higher levels of consciousness due to the greater information capacities needed to attain high mutual information. This outcome is consistent with the common intuition that conscious levels are often associated with the degree of complexity of a system.

@@

    	    %% not monotonic
    		\linelabel{line:non-monotonic}It is important to note that NTIC can be a non-monotonic function of the scale of coarse-graining. At the finest scale we consider the whole universe $X$ as the process $Y$. Then, since $Y$ is a coarse-graining of $S$ we have $Y=S=X$. In this case the environment $E$ corresponding to the universe seen as a system is the constant coarse-graining\footnote{Recall that, for a system with state space $\S$ the environment state space $\E$ must be such that $\X=\S \times \E$. If $\S =\X$ then we need $\E$ with $\X \times \E=\X$ such that $\E$ must be a singleton set. All coarse-grainings mapping $\X$ to a singleton set are constant over $\X$.} and therefore the mutual information $I(E_t;Y_{t+1})$ and the transfer entropy $I(Y_{t+1};E_t|Y_t)$ are zero. The NTIC of the universe with respect to its environment is then zero, and $X$ can never be a C-process.

    		If we now increase the scale of $Y$, this allows $S$ to also reduce in scale and therefore $E$ can become more and more fine-grained. This means that the mutual information $I(E_t;Y_{t+1})$ between $E$ and $Y$ can at least potentially become positive. Up to the point where $E$ accounts for half of the bits of $X$ and $S$ for the other half the upper bound of the mutual information $I(E_t;Y_{t+1})$ achieved when $Y=S$ increases. Refining $E$ even further again leads to a reduction of the upper bound of $I(E_t;Y_{t+1})$.
            % ==================================== %
            %    WHY DO WE NEED the FUCKING $Y$    %
            % ==================================== %
            % 2020/04/07 01:57
    		% Why does Definition 2 have a fucking $Y$? It seems the whole thing could be defined wihtout referring to the underlying $S$ and directly use $Y$'s own environment (together with which it corresponds to $X$). We would always have a refinement of $Y$ that could be called $S$ to gether with its environment. But this is alwyas the case so it doesn't restrict the definition of $Y$ at all. (Especially if we can just choose $S=Y$). The only thing that makes the current definition of $Y$ more general is that it may be that $Y$ is only closed with respect to a coarsening of its own environment and not to its own environment itself. This is a waekening of the restriction compared to requiring that $Y$ be NTIC with respect to its own environment and IC with respect to $X$. Actually no, since $Y$ is IC with respect $X$ it is IC with respect to everything including to any coarsening of its own environment. Since the mutual information between $Y$ and its environment is greater or equal to the mutual information between $Y$ and a coarsening of its environment $Y$ must also be NTIC with respect to its own environment. 
    		% ==================================== %

    		At the other extreme, when $E=X$ the system state space must be the singleton set and NTIC from $E$ to $Y$ must again be zero. Therefore, processes at intermediate scales of coarse-graining can form higher degrees of NTIC than those at the most microscopic or macroscopic scales (Fig.~\ref{fig:LevelOfConsciousness}). \ac{OurTheory} suggests that human consciousness occurs at a scale of coarse-graining where high NTIC is formed within the neural system. \linelabel{line:non-monotonic_end}

    		\begin{figure}[H]				
        		\includegraphics[width=\textwidth]{Submission/Frontiers/nonMono.pdf}
        		\caption{A non-monotonic relationship between the scale of coarse-graining and level of consciousness.}
        		\label{fig:LevelOfConsciousness}
    		\end{figure}

		\subsection{Conscious Contents Corresponding to States of an NTIC Process}\label{sec:cc}
    	    % richness 
    		\ac{OurTheory} proposes that conscious contents correspond to the states of NTIC processes (Eq.~\ref{eq:cContent}). This implies that the size of the state space of an NTIC process is associated with the richness of the conscious contents that the process can potentially have. Accordingly, a complex NTIC process with a high dimensional state space can have richer conscious experience than a simple NTIC process. This outcome is consistent with the intuition that the richness of conscious contents is associated with the complexity of a system.

    		% CC doesn't include all microscopic states
    		Informational closure can happen between scales of coarse-graining within a single system. Thus, a macroscopic NTIC process can be ignorant of its microscopic states. \ac{OurTheory} argues that human conscious contents do not reflect cellular scale activity because the conscious process which corresponds to a macroscopic NTIC process is informationally closed to the cellular scale in the human neural system. Further more, since NTIC processes are informationally closed, each of them can be considered as a reality. When the information flow from its microscopic processes (and from the environment) to it is zero (Eq.~\ref{eq:informationflow2}), the future states of the process can be entirely self-determined by its past states.

    		\linelabel{line:encode}Importantly, in most realistic cases, NTIC processes internalise the environmental dynamics in its states (see Sec. \ref{sec:causality} and also \cite{BERTSCHINGER.2006}). This suggests that an NTIC process can be considered as a process that models the environmental dynamics. This implication fits well with several theories of consciousness (for example, world simulation metaphor ~\citep{revonsuo2006inner}). Note that \ac{OurTheory} does not assume that generative models are necessary for consciousness. The implication is a natural result of processes with NTIC.
